# What is AnVIL?

[AnVIL](https://anvilproject.org/) is NHGRI's Genomic Data Science Analysis, Visualization, and Informatics Lab-Space. AnVIL inverts the traditional model of genomic data science by **bringing analysis to the data**, rather than downloading data for local analysis. By providing a secure, unified environment for data management and compute, AnVIL eliminates the need for data movement, allows for active threat detection and monitoring, and provides elastic, shared computing resources that can be acquired as needed.

<div style="display: flex; justify-content: center;">
<iframe width="560" height="315" src="https://www.youtube.com/embed/XC5qzj-yZb8?si=eq_Z0QfgH2LLAeGF" title="YouTube video player" frameborder="0" allow="accelerometer; autoplay; clipboard-write; encrypted-media; gyroscope; picture-in-picture; web-share" referrerpolicy="strict-origin-when-cross-origin" allowfullscreen></iframe>
</div>

AnVIL powers a wide range of projects, from [building a more complete human genome](https://doi.org/10.1126/science.abl3533) to supporting clinical trials to [bringing bioinformatics research to undergraduate classrooms](https://www.nature.com/articles/s41588-025-02442-5).

```{r, echo=FALSE, out.width="80%", fig.alt="Examples of high-profile publications powered by AnVIL. In particular work from the Telomere-to-Telomere project is highlighted.", fig.align="center"}
ottrpal::include_slide( "https://docs.google.com/presentation/d/1vJebvBQH_U80gDBvj_1d0USMnhaB3jiavWTLQDhgRuw/edit?slide=id.g3faae1d5063_0_971#slide=id.g3faae1d5063_0_971" )
```

These efforts are made possible by AnVIL’s ability to provide computational power and tools exactly when and where they’re needed.

So, what exactly is AnVIL? AnVIL comprises 3 pillars that together enable secure, robust, collaborative genomic data science: **data**, **compute**, and **community**.

```{r, echo=FALSE, out.width="80%", fig.alt="Diagram describing the 3 components of AnVIL: data, compute, and community.", fig.align="center"}
ottrpal::include_slide( "https://docs.google.com/presentation/d/1vJebvBQH_U80gDBvj_1d0USMnhaB3jiavWTLQDhgRuw/edit?slide=id.g409dedcc563_0_3#slide=id.g409dedcc563_0_3" )
```


## Data

AnVIL provides secure access to >5 petabytes of data, spanning hundreds of datasets and >490,000 participants.

```{r, echo=FALSE, out.width="80%", fig.alt="Graph showing the growth of AnVIL data over time, as well as a list of consortia that make data available through AnVIL.", fig.align="center"}
ottrpal::include_slide( "https://docs.google.com/presentation/d/11uxsxtnT0fIXlbXqOOglf2Yn5ciiuZuSJVAY3Mn9i18/edit?slide=id.g3f5e12e3860_0_862#slide=id.g3f5e12e3860_0_862" )
```

- **High value datasets**: AnVIL hosts many [consortia](https://anvilproject.org/consortia) datasets, including eMERGE, ALS, and GTEx.
- **Discoverable**: Through the [AnVIL Data Explorer](https://explore.anvilproject.org/datasets) you can browse or search for datasets based on many different facets, such as Diagnosis or Anatomical Site, as well as Consent Group or File Type.
- **Secure**: The AnVIL platform comes with built-in data access controls to help researchers ensure they maintain appropriate security measures while working with controlled-access data.
  - AnVIL's datasets and compute infrastructure are secured in accordance with the industry best practices, the NIST 800-53 Moderate security controls following the FedRAMP standard.
  - AnVIL is recognized by the NIH as a Controlled-Access Data Repository (CADR) that meets NIH’s Required Security and Operational Standards for NIH Controlled-Access Data Repositories (NOT-OD-25-159).
  - Learn more about [AnVIL’s Platform and Data Security](https://anvilproject.org/overview/security)


## Compute

## Community
